---
title: "aLBI - A Simple R Package for Estimating Length-Based Indicators and Fish Stock Assessment from Length Frequency Data"
author: "Ataher Ali"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simple R Package for Fish Stock Assessment}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Simple R Package for Fish Stock Assessment

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = TRUE # Ensure caching is appropriately set
)

options(repos = c(CRAN = "https://cran.rstudio.com"))
```


Introduction
This vignette provides an overview of the aLBI package...

Installation
You can install the package from GitHub using devtools:

```{r installing_package }
# Install devtools if you haven't already
if (!requireNamespace("devtools", quietly = TRUE)) {
  install.packages("devtools")
}

```

Detach or remove the package if previously attached
```{r package_management}
# Unload the aLBI package if it is already loaded
if ("package:aLBI" %in% search()) {
  detach("package:aLBI", unload = TRUE)
}

# Remove the aLBI package if it is installed
if ("aLBI" %in% rownames(installed.packages())) {
  remove.packages("aLBI")
}

```

Install and Load the Package in your R session

```{r calling_library}
# Install the Fish StockAssessment package
devtools::install_github("Ataher76/aLBI")

# Load the aLBI package
library(aLBI)

```


Data Preparation
Prepare your data in a specific format before using the functions. Ensure the data frame has two columns: Length and Frequency. Here's an example of how to load and prepare your data using the readxl package:

```{r data_preparation}
library(readxl)
# Load your length-frequency data from the system file
lenfreq_path <- system.file("exdata", "LC.xlsx", package = "aLBI")
print(lenfreq_path)  # Check the generated path

if (lenfreq_path == "") {
  stop("The required file LC.xlsx is missing. Please check the inst/extdata directory.")
}

# load the lenght frequency data
lenfreq_data <- readxl::read_excel(lenfreq_path)
print(lenfreq_data)  # check the data 
# replace with your data directory
```

Function: FishPar
The FishPar function estimates biological parameters such as Lmax, Linf, Lmat, and Lopt using bootstrapping resampling method.

Arguments

data: A data frame containing Length and Frequency columns.
resample: Number of bootstrap resamples. Default is 1000.
progress: Logical value indicating whether to show progress. Default is TRUE.

Example

```{r FishPar_Example}
# Running the FishPar function
results <- FishPar(data = lenfreq_data, resample = 1000, progress = FALSE)

# Viewing the results
results$estimated_length_par
results$estimated_froese_par
results$estimated_freq_par
results$forese_ind_vs_target
results$LM_ratio
results$Pobj

```

Output
The function returns a list with the following components:

estimated_length_par: Data frame of estimated length parameters with confidence intervals.
estimated_froese_par: Data frame of estimated Froese indicators.
estimated_freq_par: Data frame of frequency parameters.
forese_ind_vs_target: Data frame comparing Froese indicators with targets.
LM_ratio: Length at maturity ratio.
Pobj: Objective percentage combining Pmat, Popt, and Pmega.


Function: FishSS
The FishSS function evaluates stock status using criteria based on the estimated parameters.

Arguments
data: A data frame of the stock status according to Cope and Punt (2009).
LM_ratio: Length at maturity ratio from FishPar function.
Pobj: Objective percentage from FishPar function.
Pmat: Percentage of mature fish.
Popt: Percentage of optimal fish.

Example

```{r FishSS_Example}
# Load the stock status criteria data
cpdata_path <- system.file("exdata", "cpdata.xlsx", package = "aLBI")
print(cpdata_path) #check if the path exist

if (cpdata_path == "") {
  stop("The required file cpdata.xlsx is missing. Please check the inst/extdata directory.")
}
# loading the cope and punt table
cpdata <- readxl::read_excel(cpdata_path)
print(cpdata)

# Running the FishSS function
stock_status <- FishSS(data = cpdata,
                       LM_ratio = results$LM_ratio,
                       Pobj = results$Pobj,
                       Pmat = results$estimated_froese_par[1, 2],
                       Popt = results$estimated_froese_par[2, 2])

# Viewing the stock status
stock_status

```





Output
The function returns a named vector with TSB40 and LSB25 values indicating stock status.

Conclusion
The Fish Stock Assessment package provides a robust framework for estimating biological parameters and assessing fish stock status. By following the steps outlined in this vignette, you can effectively utilize this package for your fish stock assessment needs.

Additional Information
Contact
For any questions or issues, please contact the package maintainer:

Name: Ataher Ali
Email: ataher.cu.ms@gmail.com


Acknowledgements
I would like to thank the contributors and the community for their support and feedback.

This vignette provides a comprehensive guide to using the Assessment package. By following these instructions, users can effectively conduct fish stock assessments and contribute to sustainable fishery management practices.

